Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-05-23 article/02_Materials_and_Methods.Rmd
a hexagonal region grid?
R package for mantel test
adjusted figure caption accordingly

11. Defining macro-regions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data. The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age [@fokkensIntroductionBronzeAge2013]. The regions have a radius of 240km and overlap slightly (Figure \ref{map_regions} A). That means some graves are counted towards not only one but two regions. The creation of these macro-regions is necessary to establish discrete observation units, but they are also the source of methodological problems (see section \ref{discussion}).

@@

To test for correlation between the distances based on the variables *burial type* and *burial construction*, the time frame of interest was divided into sections of 200 years. For each of the $1400/200 = 7$ section and each of the two variables, a matrix of the mean SED over time was calculated. The two matrices of each section were tested for correlation with the *Mantel test*, which uses permutations of potentially correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967]. Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:dactb} C).

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
\caption{\label{correlationtimeseries}Development of correlation between different measures of cultural and spatial distance. The results for the correlation of cultural and spatial distance (Table \ref{tab:dacta}) are displayed with a solid (\textit{burial type}) and a dashed (\textit{burial construction}) line. The results for the correlation of the two cultural distances (Table \ref{tab:dactb}) are displayed with a strong dotted line \textit{without} the correction for spatial distance (only here: Pearson correlation coefficient), and with a weak dotted line for the result \textit{with} the aforementioned correction. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
\end{figure}

>>>>

11. Defining macro-regions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data. The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age [@fokkensIntroductionBronzeAge2013]. The regions have a radius of 240km and overlap slightly (Figure \ref{map_regions} A). That means some graves are counted towards not only one but two regions. A hexagonal grid layout could be considered for future analysis to avoid this overlap. The creation of these macro-regions is necessary to establish discrete observation units, but they are also the source of methodological problems (see section \ref{discussion}).

@@

To test for correlation between the distances based on the variables *burial type* and *burial construction*, the time frame of interest was divided into sections of 200 years. For each of the $1400/200 = 7$ section and each of the two variables, a matrix of the mean SED over time was calculated. The two matrices of each section were tested for correlation with the *Mantel test*, which uses permutations of potentially correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967]. The Mantel test was calculated with the function `mantel()` in the R package *ecodist* v.2.0.1 [@ecodistrpackage2007]. Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:dactb} C).

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
\caption{\label{correlationtimeseries}Development of correlation between different measures of cultural and spatial distance. The results for the correlation of cultural and spatial distance (Table \ref{tab:dacta}) are displayed with a solid (\textit{burial type}) and a dashed (\textit{burial construction}) line. The results for the correlation of the two cultural distances (Table \ref{tab:dactb}) are displayed with a strong dotted line \textit{without} the correction for spatial distance (only here: Pearson correlation coefficient), and with a weak dotted line for the result \textit{with} the aforementioned correction. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed. Statistically significant correlation (by p-Value) is indicated as such.}
\end{figure}
